### 2021-08-02 IITD_Manuscript.tex
Even clearer instructions regarding naming added.

% arara: pdflatex
% arara: biber
% arara: pdflatex

% I am sorry to have to point this out, but you cannot make any changes to the PREAMBLE of this document. It is intentionally designed to keep things generic enough so that this manuscript can be subsequently reformatted to specific journal templates/class files with minimal fuss. For initial submissions, most journals will need just the PDF file (this format) for first round of review. If there is a package you absolutely need to add, you can, but only after ensuring that you cannot live without it. Places in the preamble where you are allowed to make changes are explicitly marked with comments.

\documentclass[12 pt]{article}
% PREAMBLE starts here.

\usepackage[margin=1in,footskip=0.25in]{geometry}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{textcomp}
\usepackage{gensymb}
\usepackage{physics}
\doublespacing
\usepackage{subfigure}
\usepackage{float}
%\usepackage{caption}
\usepackage[version=4]{mhchem}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{romannum}

\usepackage[backend=biber, isbn=false, url=false, uniquename=init, terseinits=true, hyperref=true, doi=false, style=numeric, defernumbers=false, autocite=plain, sorting=none]{biblatex}
\addbibresource{publishing.bib} % Replace this with the name of your .bib file.
\usepackage[separate-uncertainty = true,multi-part-units=single]{siunitx}
% \renewcommand{\finentrypunct}{
%   \addperiod\space
%   (Cited \arabic{citecounter}~time\ifnumequal{\value{citecounter}}{1}{}{s})%
%  }
\usepackage{hyperref}
%\graphicspath{{./graphics/}} % Add this if needed. Replace graphics with Figures if that is the name of your folder.

\setlength{\marginparwidth}{2cm}\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\addcitation}[1]{\todo[inline,color=green!40]{\textbf{Add citation: }#1}}
\newcommand{\adddiscussion}[1]{\todo[inline,color=blue!40!green]{\textbf{Add discussion: }#1}}
\newcommand{\adddata}[1]{\todo[inline,color=blue!60!green]{\textbf{Add data: }#1}}
\newcommand{\explain}[1]{\todo[inline,color=red!40!yellow]{\textbf{Explain: }#1}}
\newcommand{\reviewthis}[1]{\todo[color=yellow,inline]{\textbf{Review: }#1}}
\newcommand{\response}[3]{\todo[color=yellow!90!red,inline]{\textbf{Response to Reviewer #1 (comment #2): }#3}}
\newcommand{\rework}[1]{\todo[color=orange,inline]{\textbf{Rework: }#1}}
\newcommand{\incorrect}[1]{\todo[inline,color=red!70]{\textbf{Incorrect: }#1}}
\newcommand{\rephrase}[1]{\todo[inline,color=blue!40]{\textbf{Rephrase: }#1}}
\newcommand{\warning}[1]{\todo[color=red,inline]{\textbf{Warning: }#1}}
\newcommand{\modifyfigure}[1]{\todo[inline,color=yellow!80!blue]{\textbf{Modify figure: }#1}}
\newcommand{\pleaseconfirm}[1]{\todo[inline,color=red!80!blue]{\textbf{Please confirm: }#1}}
\newcommand{\query}[1]{\todo[color=gray,inline]{\textbf{Query: }#1}}
\newcommand{\missingacknowledgment}[1]{\todo[inline,color=red!80]{\textbf{Missing acknowledgment: }#1}}
\newcommand{\restructuredocument}[1]{\todo[inline,color=blue!80]{\textbf{Restructure document: }#1}}
\newcommand{\nice}[1]{\todo[color=green,inline]{\textbf{Nice: }#1}}
\newcommand{\done}[1]{\todo[inline,color=blue!10]{\textbf{Done: }#1}}
\newcommand{\reply}[1]{\todo[color=yellow!90!red,inline]{\textbf{Reply: }#1}}

\doublespacing

% In the title, author and affil directives before, enter your details. Of course, the number of authors and affils can vary - comment out whatever you are not using/add more if needed.

\title{Preparing a manuscript for a peer-reviewed journal the FMDL way}
\author[1]{\small First Author} % ORCID ID of the first author
\author[2]{\small Second Author}  % ORCID ID of the second author
\author[1]{\small Third Author}  % ORCID ID of the third author
\author[1]{\small Postdoc Scholar}  % ORCID ID of the postdoc involved
\author[2]{\small Collaborating PI}  % ORCID ID of the collaborating PI
\author[1]{\small Madhusudan Singh\thanks{msingh@ee.iitd.ac.in}}  % ORCID ID: 0000-0001-5430-1405
\affil[1]{\footnotesize Functional Materials \& Devices Laboratory, Department of Electrical Engineering, IIT Delhi, New Delhi, NCT, India, 110 016.} % Our lab - will remain untouched.
\affil[2]{\footnotesize Department of Chemistry, IIT Delhi, New Delhi, NCT, India, 110 016.} % Change/comment out as needed

\date{}

% PREAMBLE ends here

\begin{document}
\maketitle
\begin{abstract}
This is like a conference abstract, but shorter. No use of acronymns allowed. Briefly (one sentence), motivate the manuscript. Then say what you did. Finally, summarize the main result, including salient numbers from data. This will be the last part of the paper you write just before final grammar check. This will be the \textbf{\Huge \textcolor{red}{seventh}} part of the paper you write.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Read through the entire PDF. This will tell you in which order to write the paper with cardinal order defined in bold, large red letters.

This will be the \textbf{\Huge \textcolor{red}{sixth}} part of the paper you write.

Before you do anything here, you need to complete pre-writing:

\begin{enumerate}
\item Decide what is the story. Decide what are the figures.
\item What are sections needed? What are the key citations?
\item \textbf{You have not written a word yet at this stage. Just a sketch of the article on a piece of paper.}
\item Figures. Use .eps for vector graphics. Use .tiff for scan data.
\item Shared Zotero collection. Add all the key citations. Generate a .bib file.
\item Go for a walk. Think about your competition.
\item Prepare a competition table. List relevant parameters. Add needed references to Zotero collection.
\item Scribble out a flowchart for the argument flow.
\item Decide what are competing/contradictory references that a reviewer will ask about. Add them to Zotero.
\item Go for another walk. Decide if you can address any of the weaknesses in the paper now and how much time it will take. If you can do these quickly, go back to the step on figures (add/modify).
\end{enumerate}

The moment you complete the pre-writing process, clear a full day from all distractions (social media, email, phone, etc), and write the entire paper in 6 hours. This will take 2.5-3 hours if you are putting together a Letters type manuscript. Your actual numbers in practice will be about 15\% higher than these estimates for your first paper, with the measure improving as you gain more experience.

If you have more than one actively writing co-author, you both need to first share the Zotero collection. Then, depending on your preference, you can open a document on Google Docs, and set up a Zoom call while you co-write parts of the paper together in full markup simultaneously. Alternatively, you can push your part to GitHub and inform the other student/postdoc to do a git pull and proceed. If we are collaborating with a non-LaTeX using group, the Google Doc method might work better. The backup of using pandoc to convert back and forth always exists but that is more error prone.

First step in the paper is to proceed to the results section. Once you come back to this section do the following:

\begin{itemize}
\item Paragraph one should prove the importance of the topic of your study with citations to recent (last 3-5 years) literature, especially review articles.
\item Pararaph two should clearly establish the need statement for the study by clearly defining the open problem that needs addressing.
\item Paragraph three should very briefly tell the audience what you have done (abstract style) and which approach you have followed (if that is useful to include).
\end{itemize}

Once you have done the above, and also written up an abstract, it is now time to do the following:

\begin{enumerate}
\item Cutting: Prune your sentences to increase information density. Grammarly will also help, but the first part must be done by you.
\item Check for English\cite{burchfield:1998,strunk:2000}, and run it thoroughly through Grammarly.
\item Create a github repo. Under settings of the new repo, under notifications, add your email address, and mine. What this will do is that as soon as some changes are pushed, everyone gets notified. GitHub currently allows only two email addresses.
\item Commit the first verson and push.
\item We go back and forth. The more carefully you have written the first draft, the faster this process should go.
\item Format it for the target journal identified by both of us. At that point, send the draft out to all co-authors for comments. When a collaboration is involved, I will do this myself.
\item Run a Turnitin check - remember - no repositories.
\end{enumerate}

\section{Methods}
This will be the \textbf{\Huge \textcolor{red}{first}} part of the paper you write.

In this section, clearly describe how you did your work. Create subsections for materials analysis, device fabrication, device characterization etc. The description should tersely state (within parenthesis) what tools were used, along with OEM names. In case of software/computational work, you must clearly state the algorithm used, along with any software specialized tools. Do NOT mention LabVIEW, Matlab, Python, or any such general purpose tools - that looks amateurish. Any protocols used must be cited. Do not repeat yourself - if you have previously published with a given method, cite your previous work instead of writing everything again.

\section{Results}
\label{sec:results}

This will be the \textbf{\Huge \textcolor{red}{second}} part of the paper you write.

First insert figures (Fig.~(\ref{fig:cv})) and tables. Make sure that your figures use appropriate colors, fonts, etc. \cite{mcnames:2006}. Your sentences describing your data scientifically must appear in text, and must never use reference to figures or tables except inside parenthesis. Why? When we talk to each other, we do not speak out references, we speak out ideas. In that sense, cross references and citations are ``underspeak'' that are present in citations, or in parenthesis. This has the additional merit of saving you writing space.

\begin{figure}
\centering
\subfigure[Measured capacitance as a function of area.]{\label{fig:ca}\includegraphics[width=88mm,keepaspectratio]{example-image-a}}
\subfigure[C(f) for different biases.]{\label{fig:cfv}\includegraphics[width=88mm,keepaspectratio]{example-image-b}}
\caption{(a) Use clear images, preferably in .eps form for anything other than scan data, and (b) .png form for scan data. The width of the figure is chosen to be 88mm since that is the typical width of a journal column in two-column format. Your figures must be clear to read easily at this size. If your figure is too wide to fit in this width, you can use the figure* environment, but that is done only in very rare cases where the figure is really complex.}
\label{fig:cv}
\end{figure}

The label in Fig.~(\ref{fig:cv}) is an example of a cross-reference. You \textbf{never} use absolute references in a document (like \verb|Fig.~1|). The reason for this ought to be obvious - while you author a text, things can move around, and you do not want to keep track of what moves where. \LaTeX\ is supposed to take care of that - not you. You can make mistakes, and mistakes can be costly.

An equation can be similarly labelled:

\begin{align}
  E^{2} = p^{2} c^{2} + m_{0}^{2} c^{4} \label{eqn:relativistic4vector}
\end{align}

Not all equations need labels like Eq.~(\ref{eqn:relativistic4vector}) of course:

\begin{align}
  E & = h\nu \nonumber \\
  & \equiv \frac{h c}{\lambda} \label{eqn:plancksrelationship}
\end{align}

In labeling (and referencing equations), be conservative. Label only those specific (and few) equations that you are actually going to use. Remember - any text can be made to look more inaccessible by involving more math. You should not show all steps in a paper, just the important ones that help you make your point. Needless to add, all your analytical calculations should be cross-checked with Maxima or Mathematica before it gets on your manuscript draft. No, you need to do that now. A neat trick in Mathematica is the use of \verb|TeXForm[]| function to output \LaTeX\ formatted code for the math. This may save you a lot of time, and transcription errors.

Tables can be similarly cross-referenced as shown in Table~\ref{tbl:ratio}.

\begin{table}
  \caption{An example table}
  \label{tbl:ratio}
  \begin{center}
    \begin{tabular}{cccccc}
      x  & Thickness & & \multicolumn{3}{c}{Composition: EDX and (XPS)} \\
      \hline \\
         & (nm) & & K/A & Na/A & Nb/A \\
      \hline \\
      0.3 & 62 $\pm$ 4 & & 0.365 $\pm 2\%$ & 0.634 $\pm 2\%$ & 1.036 $\pm 2\%$ \\
         & & & (0.3375) & (0.6624) & (1.028) \\
      \hline \\
      0.5 & 70 $\pm$ 2 & & 0.510 $\pm 2\%$ & 0.489 $\pm 2\%$ & 0.937 $\pm 2\%$ \\
         & & & (0.5091) & (0.4910) & (1.037) \\
      \hline \\
      0.7 & 68 $\pm$ 2 & & 0.718 $\pm 2\%$ & 0.281 $\pm 2\%$ & 0.928 $\pm 2\%$ \\
         & & & (0.7682) & (0.2317) & (1.153) \\
      \hline \\
    \end{tabular}
  \end{center}
\end{table}

Needless to add, you need to pick labels that reflect the contents of the object being referenced.

Notice some ground rules for labeling: sec for sections, fig for figures, tbl for tables. This convention, that I follow in my documents, helps distinguish between the objects that are being referenced. It is for your own benefit of course.

\section{Discussion}
\label{sec:discussion}

This will be the \textbf{\Huge \textcolor{red}{fourth}} part of the paper you write. I have intentionally split it up because even in cases when this is not a separate section in your manuscript, you should write it after a break (writing acknowledgments) - there is a difference between a) what the data are, and b) what your data mean, in light of literature.

This section is often combined with Sec.~(\ref{sec:results}). Sometimes, it is important to have it separate, especially when a fair bit of data analysis and interpretation is involved, and you need to place the discussion of the results in context of the existing literature. Whether to split or not is a personal choice, but is largely driven by how much do your results need to be compared and/or contrasted with literature. Very often, a comparison table between this work, and other studies will be placed here.

This is also the place where you should criticize certain shortcomings of your own work. It is important to put that here to disarm overly critical reviewers.

\begin{table}
  \caption{An example table}
  \label{tbl:ratio}
  \begin{center}
    \begin{tabular}{cccccc}
      x  & Thickness & & \multicolumn{3}{c}{Composition: EDX and (XPS)} \\
      \hline \\
         & (nm) & & K/A & Na/A & Nb/A \\
      \hline \\
      0.3 & 62 $\pm$ 4 & & 0.365 $\pm 2\%$ & 0.634 $\pm 2\%$ & 1.036 $\pm 2\%$ \\
         & & & (0.3375) & (0.6624) & (1.028) \\
      \hline \\
      0.5 & 70 $\pm$ 2 & & 0.510 $\pm 2\%$ & 0.489 $\pm 2\%$ & 0.937 $\pm 2\%$ \\
         & & & (0.5091) & (0.4910) & (1.037) \\
      \hline \\
      0.7 & 68 $\pm$ 2 & & 0.718 $\pm 2\%$ & 0.281 $\pm 2\%$ & 0.928 $\pm 2\%$ \\
         & & & (0.7682) & (0.2317) & (1.153) \\
      \hline \\
    \end{tabular}
  \end{center}
\end{table}

\subsection{Naming things}

There is an underlying philosophy to the way we do manuscripts in the lab - regardless of what it is, the name of the repo (already addressed in the README file), file, section, figure file, label to a figure, label to an equation, label to a table, etc. \textbf{\Huge \textcolor{red}{must reflect the contents of whatever is being referenced or named}}. Hopefully, Table~(\ref{tbl:namingthings}) will serve as a useful rolodex - it is not complete, but it is strongly indicative.

\begin{center}
  \begin{longtable}{|p{3cm}|p{5cm}|p{3.5cm}|p{2.5cm}|}
    \caption{Naming things. Mandatory parts of the string in the name/label are in blue color. DOCNAME refers to the usage in the README file of this repo.} \label{tbl:namingthings}\\
    \hline
    \textbf{Entity}  & \textbf{Purpose of the entity} &  \textbf{Good usage} & \textbf{Examples of poor choices} \\\hline
    Label to an equation & To label an equation so that it can be used in a cross-reference & \textcolor{blue}{eqn:}energyofparticle & Eq12, importantequation \\\hline
    Label to a figure & To label a figure so that it can be used in a cross-reference & \textcolor{blue}{fig:}JVdata & Fig1 \\\hline
    Label to a table & To label a table so that it can be used in a cross-reference & \textcolor{blue}{tbl:}ratio & Table1, reallyimportantsummary \\\hline
    Label to a section & To label a section in the document so that it can be used in a cross-reference & \textcolor{blue}{sec:}discussion & Sec1 \\\hline
    Name of a figure file & To name a figure file that can be subsequently called in includegraphics markup & JVdata.eps & Fig1b.eps, firstfigure.eps\\\hline
    Name of the bibliography file & To name the .bib file that you will be using in this document & \textcolor{blue}{DOCNAME.bib} & references.bib \\\hline
  \end{longtable}
\end{center}

It should be quite obvious why we follow the naming convention the way we do. If you choose to violate instructions provided in this document and the overall repo, please do us both the favor of not publishing. I do not care how good your science is, but if you cannot be bothered to use rational, common sense-based, \textbf{consistent and systematic} methods to communicate your science, you: a) are not a good scientist, and b) your communicated science will not make an impact anywhere. Good scientists are systematic, cautious, thorough, sceptical, thoughtful, and organized in a manner that makes machines look human by comparison. Yes, OCD is an occupational hazard in our business. Careful people show care and forethought in everything they can.

\section{Conclusions}
\label{sec:conclusions}

This will be the \textbf{\Huge \textcolor{red}{fifth}} part of the paper you write. This serves as a precis of your discussion, but in terms of more pithy statements, and you should highlight what the results mean for the field. This is also the place where you talk about future work.

\section*{Acknowledgments}

This will be the \textbf{\Huge \textcolor{red}{third}} part of the paper you write. Immediately after you write about results, it should be easy to remember who helped you get those results.

FA and TA (PhD fellowships) and PS (postdoc fellowship) acknowledge support from Ministry of Human Resource \& Development (MHRD). SA acknowledges support from University Grants Commission. CP acknowledges partial support from grant XYZ from Department of Science \& Technology. PS and MS acknowledges support from grant ABC from Department of Science \& Technology. All authors acknowledge facility access to Central Research Facility (CRF) and Nanoscale Research Facility (NRF, NNetra program) at IIT Delhi. Authors acknowledge technical assistance from Mr. Did Occasional Measurements/Process Runs of CRF.

Authors FA, TA, PS and MS declare competing interest in the form of an Indian patent application (2021345678).

\printbibliography

\end{document}