### 2022-12-11 common/preamble.tex

\usepackage[margin=1in,footskip=0.25in]{geometry}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{textcomp}
\usepackage{gensymb}
\usepackage{array}
\doublespacing
\usepackage{subfigure}
\usepackage{float}
%\usepackage{caption}
\usepackage[version=4]{mhchem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{authblk}
\usepackage{romannum}
\usepackage{caption} % For Table of Contents entry
\usepackage{lineno}
\usepackage{booktabs}

\usepackage[backend=biber, isbn=false, url=false, uniquename=init, terseinits=true, hyperref=true, doi=false, style=numeric, defernumbers=false, autocite=plain, sorting=none]{biblatex}
\usepackage[separate-uncertainty=true,multi-part-units=single,per-mode=symbol,range-units=single,range-phrase=\text{-}]{siunitx}
\DeclareSIUnit{\Molar}{M}
\DeclareSIUnit\gauss{G}
\DeclareSIUnit\sq{\ensuremath{\Box}}
\DeclareSIUnit\degc{\unit{\degreeCelsius}}
\usepackage{hyperref}